Uncovering Distributed System Bugs\\ during Testing (not Production!)

### 2015-08-18 paper.tex
fix in additional authors
added a picture depicting the high level architecture of xstore vnext
merged bg with overview
added some cites
Starting to describe P# async integration for migration.

\numberofauthors{7} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor Pantazis Deligiannis\\
       \affaddr{Imperial College London}\\
       \affaddr{London, UK}\\
       \email{p.deligiannis@imperial.ac.uk}
% 2nd. author
\alignauthor John Erickson\\
       \affaddr{Microsoft}\\
       \affaddr{Redmond, WA, USA}\\
       \email{jerick@microsoft.com}
% 3rd. author
\alignauthor Cheng Huang\\
       \affaddr{Microsoft Research}\\
       \affaddr{Redmond, WA, USA}\\
       \email{cheng.huang@microsoft.com}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Matt McCutchen\\
       \affaddr{MIT}\\
       \affaddr{Cambridge, MA, USA}\\
       \email{rmccutch@mit.edu}
% 5th. author
\alignauthor Shaz Qadeer\\
       \affaddr{Microsoft Research}\\
       \affaddr{Redmond, WA, USA}\\
       \email{qadeer@microsoft.com}
% 6th. author
\alignauthor Wolfram Schulte\\
       \affaddr{Microsoft}\\
       \affaddr{Redmond, WA, USA}\\
       \email{schulte@microsoft.com}
}

\date{15 August 2015}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

How cheng was testing his code before and how after -- high level, pretty picture

%\input{overview}

\section{Background}
\label{sec:bg}

\input{background}

Dependency injection pattern

@@

% cheng to review
\subsection{Distributed Extent Management System}

@@

We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing it.  MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.  It moves all data from the old table to the new table in the background.  Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.

@@

% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.  Given a multi-page read, SpecTable can actually produce a list of all valid results.  Our correctness property is then:
\begin{quote}
For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
\end{quote}
%
\def\term#1{\emph{#1}}
We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).  Specifically, after each backend call, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.  This makes it possible to verify the correctness property as the system executes.  We have a \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.  Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.  When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.  When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.  \psharp controls the interleaving of the backend calls.  After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.  We also use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp scheduling strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.

To do...

>>>>

\numberofauthors{6} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor Pantazis Deligiannis\\
       \affaddr{Imperial College London}\\
       \affaddr{London, UK}\\
       \email{p.deligiannis@imperial.ac.uk}
% 2nd. author
\alignauthor John Erickson\\
       \affaddr{Microsoft}\\
       \affaddr{Redmond, WA, USA}\\
       \email{jerick@microsoft.com}
% 3rd. author
\alignauthor Cheng Huang\\
       \affaddr{Microsoft Research}\\
       \affaddr{Redmond, WA, USA}\\
       \email{cheng.huang@microsoft.com}
\and  % use '\and' if you need 'another row' of author names
% 4th. author
\alignauthor Matt McCutchen\\
       \affaddr{MIT}\\
       \affaddr{Cambridge, MA, USA}\\
       \email{rmccutch@mit.edu}
% 5th. author
\alignauthor Shaz Qadeer\\
       \affaddr{Microsoft Research}\\
       \affaddr{Redmond, WA, USA}\\
       \email{qadeer@microsoft.com}
% 6th. author
\alignauthor Wolfram Schulte\\
       \affaddr{Microsoft}\\
       \affaddr{Redmond, WA, USA}\\
       \email{schulte@microsoft.com}
}

\date{17 August 2015}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

\input{overview}

@@

% cheng to review
\subsection{Distributed Extent Management System}
\label{sec:cases:azurestore}

@@

We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing the data set.  MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.  It moves all data from the old table to the new table in the background.  Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.

@@

% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.  Given a multi-page read, SpecTable can actually produce a list of all valid results.  Our correctness property is then:
\begin{quote}
For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
\end{quote}
%
\def\term#1{\emph{#1}}
We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).  Specifically, after each backend call, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.  This makes it possible to verify the correctness property as the system executes.  We have a \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.  Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.  When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.  When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.  \psharp controls the interleaving of the backend calls.  After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.  We use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp behavior exploration strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.

% Draft
We wanted to implement the core MigratingTable algorithms in \csharp ``async'' code, like most of Artifact Services.  Async code is readable like traditional procedural code but is translated by the compiler to an event-driven state machine using the Task Parallel Library, gaining most of the performance benefits of that style.  By default, all event processing occurs on the .NET thread pool.

We had to arrange for the continuations to execute within the context of a single \psharp machine, which we did by installing a custom SynchronizationContext.  Then we implemented async RPC between machines in terms of message passing, using RealProxy to avoid most of the marshaling boilerplate.
